---
title: "Zip Code Smoothing"
author: "Andres Potapczynski (ap3635)"
date: "11/9/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Imports, echo=FALSE, message=FALSE, warning=FALSE}
library(rstan)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(bayesplot)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```

# Summary

The purpose of this `Rmarkdown` is to see whether it makes sense that the zip codes have different default probabilities or else all can be regularized. The first approach will be to regularized them as if they come from the same distribution. The second approach will be to see if they come from a mixture.

The proposed model is the following
$$
y_j \sim Binomial(n_j, \theta_j)
$$
and we will also try
$$
y_j \sim Poisson(n_j \theta_j)
$$
since the Poisson distribution is more suitable for rare events.

Moreover, each $\theta_j$ is assumed to come from a Beta distribution.

$$
\theta_j \sim Beta(\alpha, \beta)
$$

where the hyperparameters can be obtained via
$$
\alpha \sim Ga(2, 2)
$$
and 
$$
\beta \sim Ga(2, 2)
$$
The second approach has the following generative process [...]

# Run the analysis

## Load the data

```{r Load the data, message=FALSE, warning=FALSE}
file <-  './DBs/core.txt'
data <- read_delim(file = file, delim = '|')

# Sample the data
# pct = 1
pct = 0.1
# pct = 0.01
set.seed(seed = 42)
sample_size = round(pct * nrow(data))
sample <- sample(x = nrow(data), size = sample_size, replace = F)
data = data[sample, ]

# Change column format
data$postal_code = factor(data$postal_code)
```

Let's understand the distribution of the zip codes in the DB.
```{r Distribution of zip_codes in the data}
zip_summary = data %>% 
  # group_by(postal_code, state) %>% 
  # group_by(postal_code) %>% 
  group_by(postal_code, city, state) %>% 
  summarize(mort_no = n(), 
            y_sum = sum(y), 
            income_bar = mean(client_income), 
            factor_bar = mean(factor_employed),
            lender_bar = mean(lender_score),
            risk_bar = mean(risk_index),
            ratio_bar = mean(ratio),
            asset_val = mean(asset_market_value)) %>% 
  mutate(theta_emp = y_sum / mort_no) %>% 
  arrange(desc(mort_no))

N = zip_summary$mort_no
y = zip_summary$y_sum
M = nrow(zip_summary)
```

The distribution of the zip codes is

```{r Plot empirical dist of default per each zip code, eval=FALSE}
ggplot(data = zip_summary, mapping = aes(x=theta_emp)) +
  geom_histogram(fill='lightblue', color='black', binwidth = 0.05)
```

```{r Plot the zipcode vs probability of default, eval=TRUE}
zip_df <- zip_summary
# zip_df = zip_summary %>% filter(theta_emp > 0)
ggplot(data = zip_df, mapping = aes(x=income_bar, y=theta_emp)) +
  geom_point()
```


Now, I compile the proposed STAN model.
```{r Compile the hierarchical STAN model for 1}
sm <- stan_model('./zip_code_v01.stan')
```

## Run first approach

The data for the first model is

```{r Generate data for the first approach}
inputs = list(M=M, N=N, y=y)
model.v01 = sampling(sm, data=inputs)
```

```{r Print the results of the first approach}
print(model.v01, digits=2, pars = c('alpha', 'beta'))
```

### Look at the shrinkage of the model

```{r Extract the thetas for the first model, eval=TRUE}
sims = rstan::extract(model.v01)
theta = apply(X = sims$theta, MARGIN = 2, FUN = median)
df = data.frame(theta=theta)
g_post <- ggplot(data = df, mapping = aes(x = theta)) +
  geom_histogram(fill='lightblue', color='black', binwidth = 0.01) +
  ggtitle('Beta-Binomial Model') +
  xlim(0, 0.2) +
  ylim(0, 4500)
g_post
```

```{r Plot the regularized parameters}
sims = rstan::extract(model.v01)
theta = apply(X = sims$theta, MARGIN = 2, FUN = median)
df2 <- data.frame(income_bar=zip_summary$income_bar, theta=theta)
ggplot(data = df2, mapping = aes(x=income_bar, y=theta)) +
  geom_point() +
  ylim(0, 1)
```

### Generate data

```{r Generate y_rep for evaluation}
sims = rstan::extract(model.v01)
y_rep_max = apply(X = sims$y_rep, MARGIN = 1, FUN = max)
df_rep = data.frame(y_rep_max = y_rep_max)
ggplot(data = df_rep, mapping = aes(x = y_rep_max)) + 
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = max(zip_summary$y_sum))
```


## Run second approach

Now, I compile the proposed STAN model.
```{r Compile the hierarchical STAN model for poisson for 2}
sm_poi <- stan_model('./zip_code_v02.stan')
```

The data for the second model is

```{r Generate data for the second approach}
inputs = list(M=M, N=N, y=y)
model.v02 = sampling(sm_poi, data=inputs)
```

```{r Print the results of the second approach}
print(model.v02, digits=2, pars = c('alpha', 'beta'))
```

```{r Extract the thetas for the second model, eval=TRUE}
sims.poi = rstan::extract(model.v02)
theta = apply(X = sims.poi$theta, MARGIN = 2, FUN = median)
df = data.frame(theta=theta)
g_post.poi <- ggplot(data = df, mapping = aes(x = theta)) +
  geom_histogram(fill='lightblue', color='black', binwidth = 0.01) + 
  ggtitle('Poisson Model') +
  xlim(0, 0.2) +
  ylim(0, 4500)
gridExtra::grid.arrange(g_post, g_post.poi, nrow=1, ncol=2)
```

```{r Plot the regularized parameters for Poisson}
sims.poi = rstan::extract(model.v02)
theta = apply(X = sims.poi$theta, MARGIN = 2, FUN = median)
df2 <- data.frame(income_bar=zip_summary$income_bar, theta=theta)
ggplot(data = df2, mapping = aes(x=income_bar, y=theta)) +
  geom_point() +
  ylim(0, 1)
```

## Run clustering approach

Now, I compile the proposed STAN model.
```{r Compile the hierarchical STAN model for 3}
sm_cluster <- stan_model('./zip_code_v03.stan')
```

The data for the third model is

```{r Generate data for the third approach, eval=FALSE}
K = 2
z0 = sample(x = K, size = M, replace=TRUE)
inputs = list(K=K, M=M, N=N, y=y, a0=rep(1, 2), z=z0)
model.v03 = sampling(sm_cluster, data=inputs)
```

```{r Generate data for the clustering approach}
K = 2
z0 = sample(x = K, size = M, replace=TRUE)
inputs = list(K=K, M=M, N=N, y=y, a0=rep(1, 2), z=z0)
model.opt = optimizing(sm_cluster, data=inputs)
```

were the results of the model are

```{r Print the results of the third approach, eval=FALSE}
print(model.v03, digits=2, pars = c('alpha', 'beta', 'phi', 'theta'))
```

```{r Count the number of bad zip codes, eval=FALSE}
sims.cluster = rstan::extract(model.v03)
```


# Concluding remarks

Add the resulting $\theta_j$s to the `zip summary` table.

```{r Merge back}
theta_beta = apply(X = sims$theta, MARGIN = 2, FUN = median)
theta_poi = apply(X = sims.poi$theta, MARGIN = 2, FUN = median)
zip_summary$theta_beta = theta_beta
zip_summary$theta_poi = theta_poi
path = './DBs/zip.txt'
write_csv(x = zip_summary, path = path)
```


[...]