---
title: "Logistic Regression Take 01"
author: "Andres Potapczynski (ap3635)"
date: "11/9/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Imports, echo=FALSE, message=FALSE, warning=FALSE}
library(rstan)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(bayesplot)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```


# Results from first logistic regression

```{r Load the data, warning=FALSE, message=FALSE, echo=FALSE}
# Load provisional data set
# file <-  './DBs/core_sample.txt'
file <-  './DBs/core.txt'
data <- read_delim(file = file, delim = '|')

# Sample the data
pct = 0.1
set.seed(seed = 42)
sample_size = round(pct * nrow(data))
sample <- sample(x = nrow(data), size = sample_size, replace = F)
data = data[sample, ]

# Change column format
data$new_used = factor(x = data$new_used, levels = c('U', 'N'))
data$sex = factor(x = data$sex, levels = c(1, 2))
data$postal_code = as.integer(data$postal_code)
```

## Plot variable distributions

Now we have the following age distribution

```{r Plot age distribution, echo=FALSE}
ggplot(data = data, mapping = aes(x=age)) +
  geom_histogram(binwidth = 1, fill='lightblue', color='black') +
  xlab('Age (years)') + 
  ylab('Count') + 
  labs(title='Young adults are more prevalent in the data', 
       subtitle = 'Age distribution') + 
  theme(plot.title = element_text(size = 12, face='bold')) + 
  theme(plot.subtitle = element_text(size = 10))
```

Also, the distribution of the number of elements per zip code is the following

```{r Plot zipcode information, eval=FALSE}
zip_summary = data %>% 
  group_by(postal_code) %>% 
  summarize(mort_no = n()) %>% 
  arrange(desc(mort_no))

reversal = zip_summary %>% 
  group_by(mort_no) %>% 
  summarize(count = n())
ggplot(data = reversal, mapping = aes(x = count)) +
  geom_histogram(binwidth = 10, fill='lightblue', color='black') +
  xlab('No of mortages') + 
  ylab('Count') + 
  xlim(0, 100) +
  labs(title='...', 
       subtitle = 'Number of mortgages per zip code') + 
  theme(plot.title = element_text(size = 12, face='bold')) + 
  theme(plot.subtitle = element_text(size = 10))
```

## Run model

```{r Select the relevant columns for the analysis}
y = data$y
data_sub <- data %>% select(client_income,
                        ratio,
                        age,
                        asset_market_value,
                        lender_score,
                        factor_employed,
                        risk_index)
summary(data_sub)
```

## Evaluate the model

I compile the model to access it later and then run it in a separate code chunk.

```{r Compile the STAN model}
sm <- stan_model('./logistic_reg_v01.stan')
```


```{r Function to replay analysis}
### Debug this function
single_analysis <- function(var, data, y, sm){
  input_data = data %>% select(var)
  inputs = list(N = nrow(input_data), D=ncol(input_data), X=input_data, y=y)
  model.fit = sampling(sm, data=inputs, verbose=F)
  return(model.fit)
}
```


### For age only

```{r Compile the STAN model for age}
input_data = data_sub %>% select(age)
inputs = list(N = nrow(input_data), D=ncol(input_data), X=input_data, y=y)
```

```{r Run the STAN model for age, echo=FALSE}
model_v01 = sampling(sm, data=inputs, verbose=F)
```

The results are the following
```{r Print results of model v01 for age}
print(model_v01, digits = 2)
```

### For risk index only

```{r Compile the STAN model for risk index}
input_data = data_sub %>% select(risk_index)
inputs = list(N = nrow(input_data), D=ncol(input_data), X=input_data, y=y)
```

```{r Run the STAN model for risk index, echo=FALSE}
model_v01 = sampling(sm, data=inputs, verbose=F)
```

The results are the following
```{r Print results of model v01 for risk index}
print(model_v01, digits = 2)
```

### For asset market value only

```{r Compile the STAN model for asset market value}
input_data = data_sub %>% select(asset_market_value)
inputs = list(N = nrow(input_data), D=ncol(input_data), X=input_data, y=y)
```

```{r Run the STAN model for asset market value, echo=FALSE}
model_v01 = sampling(sm, data=inputs, verbose=F)
```

The results are the following
```{r Print results of model v01 for asset market value}
print(model_v01, digits = 2)
```

were we see that there was a lot of problem with convergence in this case.

### For client income

```{r Compile the STAN model for client income}
input_data = data_sub %>% select(client_income)
inputs = list(N = nrow(input_data), D=ncol(input_data), X=input_data, y=y)
```

```{r Run the STAN model for client income, echo=FALSE}
model_v01 = sampling(sm, data=inputs, verbose=F)
```

The results are the following
```{r Print results of model v01 for client income}
print(model_v01, digits = 2)
```

were there is convergence but, yet again, one of the variables is not important again.

### For ratio

```{r Compile the STAN model for ratio}
input_data = data_sub %>% select(ratio)
inputs = list(N = nrow(input_data), D=ncol(input_data), X=input_data, y=y)
```

```{r Run the STAN model for ratio, echo=FALSE}
model_v01 = sampling(sm, data=inputs, verbose=F)
```

The results are the following
```{r Print results of model v01 for ratio}
print(model_v01, digits = 2)
```

were, so far, this is the only variable that has shown some significance.

```{r Extract variables for ratio}
sims = rstan::extract(model_v01)
alpha_mean = apply(X = sims$alpha, MARGIN = 1, FUN = mean)
beta_mean = apply(X = sims$beta, MARGIN = 1, FUN = mean)
true_params = c(alpha_mean, beta_mean)
posterior_params = as.matrix(model_v01, pars=c('alpha', 'beta'))
# mcmc_recover_hist(posterior_params)
mcmc_hist(posterior_params, binwidth = 0.25)
```


### For lender score

```{r Compile the STAN model for lender score}
input_data = data_sub %>% select(lender_score)
inputs = list(N = nrow(input_data), D=ncol(input_data), X=input_data, y=y)
```

```{r Run the STAN model for lender score, echo=FALSE}
model_v01 = sampling(sm, data=inputs, verbose=F)
```

The results are the following
```{r Print results of model v01 for lender score}
print(model_v01, digits = 2)
```

were apparently there is a little effect.

```{r Extract variables for lender score}
sims = rstan::extract(model_v01)
alpha_mean = apply(X = sims$alpha, MARGIN = 1, FUN = mean)
beta_mean = apply(X = sims$beta, MARGIN = 1, FUN = mean)
true_params = c(alpha_mean, beta_mean)
posterior_params = as.matrix(model_v01, pars=c('alpha', 'beta'))
# mcmc_recover_hist(posterior_params)
mcmc_hist(posterior_params)
```

### For factor employed

```{r Compile the STAN model for factor employed}
input_data = data_sub %>% select(factor_employed)
inputs = list(N = nrow(input_data), D=ncol(input_data), X=input_data, y=y)
```

```{r Run the STAN model for factor employed, echo=FALSE}
model_v01 = sampling(sm, data=inputs, verbose=F)
```

The results are the following
```{r Print results of model v01 for factor employed}
print(model_v01, digits = 2)
```

were in the previous variable the beta never went past 0 but here we do have this effect.

###
