---
title: "Final_Yi"
output: html_document
date: "`r format(Sys.Date())`"
editor_options: 
  chunk_output_type: console
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model Extension: CAR & Zero Inflation 

### A linear binormial regression analogy

Lik what we did in the baseline model. We process by treating the underlying deterministic model as providing an expected default times for each city around which there will be variation due to both measurement error and simplifications. Consider the typical formulation of a linear regression, where $y_n$ is the is an observable default time, $x_n$ is a row vector of unmodeled predictors ( independent variables), $\beta$ is a coefficient vector parameter and we separate the intercep as $a$. In the city level, we assume that the number of individual records in city $i$ is $n_i$.Thus, we have the model:

$$ y_i \sim Binormial(n_{i},logit^{-1}(a + x_i\beta)) $$
As a robust prior distribution option, we take $\beta \sim cauchy(location=0,scale=2.5)$.

In model 2, we will extend this model in two ways: (1) incoperate the geographic state information (2) zero inflated model.  

### Extension one: Incoperate Geographic Information

In this project, we utilize the IAR prior for state feature. Intrisnic conditional autoregressive (IAR)  is an extension of conditional autoregressive (CAR) models, which are popular as prior distributions for spatial random effects with areal spatial data. In our model, we have a random quantity $\phi = (\phi_1,\phi_2,...,\phi_{32})$ at 32 state areal locations. In each state, we have the individual records aggregated at the city level. And each city data belong to one state. According to the Brook's Lemma, the joint the distribution of $\phi$ can be expressed as the followings:
$$\phi \sim N(0,[D_{\tau} (I-\alpha B)]^{-1})$$
In this formula, we have:

* $D = diag(m_i)$ is an $32 \times 32$ diagonal matrix with $m_i$ is the number of the neighbors for the state i

* $D_{\tau} = \tau D$  and $\tau$ is the hyperparameter in the conditional distributions of the $\phi$

* $\alpha$ is the parameter that controls spatial dependence. In IAR, we let $\alpha =1$

* $B = D^{-1} W$ is the scaled adjacency matrix. And $W$ is the adjacency matrix. ($w_{ii}=0$,$w_{ij}=1$ if the state i is a neighbor of state j , and $w_{ij}=1$ otherwise)

We can simplifies the IAR model to:

$$\phi \sim N(0,\tau (D-W)]^{-1})$$
In IAR model, we have a singular precision matrix and an improper prior distribution. However, in practice, IAR models are fit with a sum to zero constrains: $\sum_{i} \phi_i = 0 $ for each connected component of the graph. In this way, we can interpret both overall means and the component-wise means.

Through log probability accumulator, we can accure computational efficiency gains. We have:

$$log(p(\phi | \tau)) = -\frac{n}{2}log(2\pi) + \frac{1}{2}log(det^{*}(\tau(D-W))) - \frac{1}{2}\phi^{T}\tau(D-W)\phi$$
$$=-\frac{n}{2}log(2\pi) + \frac{1}{2}log(\tau^{n-k}) + \frac{1}{2}log(det^{*}(\tau(D-W))  - \frac{1}{2}\phi^{T}\tau(D-W)\phi$$

In this formula, $det^{*}(A)$ is the generlized determinant of the square matrix A defined as the product of its non-zero eigenvalues, and the k is the number of the connented component in the graph.(k=1 for our data) Dropping the additive constants, the qunantity to increment becomes:

$$  \frac{1}{2}log(\tau^{n-k}) - \frac{1}{2}\phi^{T}\tau(D-W)\phi $$

In  our model, we assume the hyperparameter $\tau \sim Gamma(shape = 2,rate = 2)$. We define the sparse_iar_lpdf function as a more efficient spare representation as the following:

```{r eval=FALSE}
functions {
  real sparse_iar_lpdf(vector phi, real tau, int[,] W_sparse, vector D_sparse, vector lambda, int S, int W_n) {
      row_vector[S] phit_D; // phi' * D
      row_vector[S] phit_W; // phi' * W
      vector[S] ldet_terms;
    
      phit_D = (phi .* D_sparse)';
      phit_W = rep_row_vector(0, S);
      for (i in 1:W_n) {
        phit_W[W_sparse[i, 1]] = phit_W[W_sparse[i, 1]] + phi[W_sparse[i, 2]];
        phit_W[W_sparse[i, 2]] = phit_W[W_sparse[i, 2]] + phi[W_sparse[i, 1]];
      }
    
      return 0.5 * ((S-1) * log(tau)
                    - tau * (phit_D * phi - (phit_W * phi)));
  }
}
```

After we get the IAR prior, we can take it into our model.For the $i$ the city in the $j$ the state, we have:

$$ y_{ij} \sim Binormial(n_{ij},logit^{-1}(a + \phi_j + x_{ij}\beta)) $$
* Reason for design the model in this way rather than hierarchical model:

1. Hierarchical extension will greatly expend the demension of the parameter space. Consequently, the estimation convergence would be much more difficult.Just focus on the binormial regression part. If we take the hierarchical extension both on the intercept and coefficient terms, there will have $32 \times 1 = 32$ intercept paramters and $32 \times 7 = 224$ coefficient parameters (total 256 parameters).For taking  the hierarchical extension on  $\beta$s, we have 1 intercept and $7 \times 32 = 224$ coefficient in the binormial regression model (total 225 parameters). But now, we only have 1 parameter for overall intercept, 32 parameter for the state level effecr and 7 parameter for the different independent variables' effects.

2. This design of model is better for interpreation and better for us to solve our research problems.Now the coefficient terms are not depend on the state prior. Thus, we can estimate the overall effect from these independent variables, like age and income. And on the state level effect, we can have the overall idea based on the estimated parameter values.

### Extension two: zero inflated model

Zero-inflated model originally provide mixture of a mixtures of a Poisson and Bernoulli probability mass function to allow more flexibility in modeling the probability of a zero outcome. Zero-inflated models, as defined by Lambert (1992), add additional probability mass to the outcome of zero. But, this extension can also be applied for other categorical distributions like binomial distribution we used in this project. 

We assume a parameter $\theta$ as the probability of drawing a zero and the probability $1-\theta$ as drawing fro mthe Binormal distribution. The prior distribution of  $\theta$ is uniform between 0 and 1, since we have no extra information about this parameter. The distribution function is thus:

$$p(y_n | \theta,a,\beta) =\begin{cases} 
\theta + (1-\theta) \times Binomial(0 |a,\beta,\phi ) & y_n =0 \\
(1-\theta) \times Binomial(0 |a,\beta,\phi )  & y_n > 0 \end{cases}$$

In stan, we estimate the model in this following ways:

```{r eval=FALSE}
for (j in 1:N_train){
    if (y[j] == 0){
      target += log_sum_exp(bernoulli_lpmf(1 | theta), bernoulli_lpmf(0 | theta) + binomial_logit_lpmf(y[j] | n_city_train[j],alpha[state_train[j]] + X_train[j,]* beta));
      }
    else{
      target += bernoulli_lpmf(0 | theta) + binomial_logit_lpmf(y[j] | n_city_train[j],alpha + X_train[j,] * beta);}
  }
```

And we predict the $y_{rep}$ in the following way:

```{r eval=FALSE}
generated quantities{
  int y_rep[N_train];
  real<lower =0,upper=1> zero_train[N_train];
  for (i in 1:N_train){
    zero_train[i] = uniform_rng(0,1);
    if (zero_train[i] < theta){
      y_rep[i] = 0;
      }
    else{
      y_rep[i] = binomial_rng(n_city_train[i],inv_logit(alpha[state_train[i]] + X_train[i,]* beta));
      }
  }
}
```


### Model Coveage testing wit fake data.

In this part, we first simulate the fake data as we assume in thie model. Then, we will check that our model works well with the data that we have simulated ourselves. In this following are the model coverage plots.








### Model result

In the following is the basic model results:

```{r eval=FALSE}

```




### Model summary

In order to determine our model performance, again we do the 5-floder cross validation. And we calculate the MSE for each training dataset with our model. And then we get the average MSE. The stan code we used to simulate the $y_hat$ is as the following:

```{r eval=FALSE}
generated quantities{
  int y_rep_cv[N_test];
  real<lower =0,upper=1> zero_test[N_test];
  for (i in 1:N_test){
    zero_test[i] = uniform_rng(0,1);
    if (zero_test[i] < theta){
      y_rep_cv[i] = 0;
      }
    else{
      y_rep_cv[i] = binomial_rng(n_city_test[i],inv_logit( alpha[state_test[i]] + X_test[i,]* beta));
      }
  }
}
```


Acrroding to the result, the baseline MSE is 128164. But for our model, the average MSE is **7583** with the standard deviation **6831**. Thus, we can say that our model have a huge improve from the baseline.





























```{r  include=FALSE}
library(knitr);library(arm);library(ggplot2)
library(MASS);library(tidyr);library(dplyr)
library(extraDistr);library(gridExtra)
library(rstan);library(bayesplot);library(loo)
library(shinystan);library(readr)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

knitr::opts_chunk$set(
	eval = FALSE,
	error = FALSE,
	message = FALSE,
	warning = FALSE,
	comment = NA,
	include = FALSE
)
options(xtable.comment = FALSE)
options(htmltools.dir.version = FALSE)
options(digits = 2)
```


```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
file <-  '../data/core.txt'
data <- read_delim(file = file, delim = '|')

# Sample the data
pct = 1
# pct = 0.1
# pct = 0.01
set.seed(seed = 42)
sample_size = round(pct * nrow(data))
sample <- sample(x = nrow(data), size = sample_size, replace = F)
data = data[sample, ]
## Selecting the relevant columns for the analysis
data_sub <- data %>% dplyr::select(
  state,
  city,
  county,
  zip,
  asset_market_value,
  mar_2_app,
  appraisal_value,
  app_2_inc,
  client_income,
  mar_2_inc,
  age,
  sex_F,
  condition_U,
  y,
  y2)

## Group data by state and define the IDs
state_summary <- data_sub %>% 
  dplyr::select(state, 
                client_income, 
                appraisal_value,
                asset_market_value) %>% 
  group_by(state) %>% 
  summarize(n_state = n(),
            income_mean_state = mean(client_income),
            appraisal_mean_state = mean(appraisal_value),
            market_mean_state = mean(asset_market_value)) %>% 
  arrange(desc(n_state)) %>% 
  ungroup()
state_summary$ID_state = seq.int(nrow(state_summary))


## Group data by city and define the IDs
city_summary <- data_sub %>% 
  dplyr::select(city, state,
                client_income,
                appraisal_value,
                asset_market_value,
                mar_2_inc,
                mar_2_app,
                app_2_inc,
                age,
                y,
                y2) %>% 
  group_by(city, state) %>% 
  summarize(n_city = n(),
            income_mean_city = mean(client_income),
            appraisal_mean_city = mean(appraisal_value),
            market_mean_city = mean(asset_market_value),
            
            mar_2_inc_mean_city = mean(mar_2_inc),
            mar_2_app_mean_city = mean(mar_2_app),
            app_2_inc_mean_city = mean(app_2_inc),
            
            age_mean_city = mean(age),
            sum_y = sum(y),
            sum_y2 = sum(y2)) %>% 
  arrange(desc(n_city)) %>% 
  ungroup()


## Merge back into data
city_summary <- city_summary %>% 
  inner_join(y = state_summary[c('ID_state', 'state')], by = 'state')
```


```{r eval=FALSE, include=FALSE}
N = length(unique(city_summary$state))
A <- matrix(0,ncol = N,nrow = N)
colnames(A) = unique(city_summary$state)
rownames(A) = unique(city_summary$state)
A['NUEVO LEON',c('ZACATECAS','SAN LUIS POTOSI','TAMAULIPAS','COAHUILA DE ZARAGOZA')] = 1
A[c('ZACATECAS','SAN LUIS POTOSI','TAMAULIPAS','COAHUILA DE ZARAGOZA'),'NUEVO LEON'] = 1
A['VERACRUZ LLAVE',c('TAMAULIPAS','SAN LUIS POTOSI','HIDALGO','PUEBLA','OAXACA','CHIAPAS','TABASCO')] = 1
A[c('TAMAULIPAS','SAN LUIS POTOSI','HIDALGO','PUEBLA','OAXACA','CHIAPAS','TABASCO'),'VERACRUZ LLAVE'] = 1
A['GUANAJUATO',c('MICHOACAN DE OCAMPO','JALISCO','SAN LUIS POTOSI','ZACATECAS','QUERETARO DE ARTEAGA')] = 1
A[c('MICHOACAN DE OCAMPO','JALISCO','SAN LUIS POTOSI','ZACATECAS','QUERETARO DE ARTEAGA'),'GUANAJUATO'] = 1
A['NAYARIT',c('SINALOA','DURANGO','ZACATECAS','JALISCO')] = 1
A[c('SINALOA','DURANGO','ZACATECAS','JALISCO'),'NAYARIT'] = 1
A['JALISCO',c('NAYARIT','ZACATECAS','COLIMA','MICHOACAN DE OCAMPO','GUANAJUATO','AGUASCALIENTES')] = 1
A[c('NAYARIT','ZACATECAS','COLIMA','MICHOACAN DE OCAMPO','GUANAJUATO','AGUASCALIENTES'),'JALISCO'] = 1
A['DISTRITO FEDERAL',c('ESTADO DE MEXICO','MORELOS')] = 1
A[c('ESTADO DE MEXICO','MORELOS'),'DISTRITO FEDERAL'] = 1
A['MORELOS',c('GUERRERO','PUEBLA','ESTADO DE MEXICO')] = 1
A[c('GUERRERO','PUEBLA','ESTADO DE MEXICO'),'MORELOS'] = 1
A['MICHOACAN DE OCAMPO',c('COAHUILA DE ZARAGOZA','ESTADO DE MEXICO','JALISCO','GUANAJUATO','GUERRERO','QUERETARO DE ARTEAGA')] = 1
A[c('COAHUILA DE ZARAGOZA','JALISCO','ESTADO DE MEXICO','GUANAJUATO','GUERRERO','QUERETARO DE ARTEAGA'),'MICHOACAN DE OCAMPO'] = 1
A['ESTADO DE MEXICO',c('HIDALGO','QUERETARO DE ARTEAGA','MICHOACAN DE OCAMPO','GUERRERO','PUEBLA','TLAXCALA','MORELOS')] = 1
A[c('HIDALGO','QUERETARO DE ARTEAGA','MICHOACAN DE OCAMPO','GUERRERO','PUEBLA','TLAXCALA','MORELOS'), 'ESTADO DE MEXICO'] = 1
A['HIDALGO',c('ESTADO DE MEXICO','QUERETARO DE ARTEAGA','SAN LUIS POTOSI','VERACRUZ LLAVE','PUEBLA','TLAXCALA')] = 1
A[c('ESTADO DE MEXICO','QUERETARO DE ARTEAGA','SAN LUIS POTOSI','VERACRUZ LLAVE','PUEBLA','TLAXCALA'),'HIDALGO'] = 1
A['ZACATECAS',c('JALISCO','NAYARIT','DURANGO','COAHUILA DE ZARAGOZA','NUEVO LEON','TAMAULIPAS','SAN LUIS POTOSI','AGUASCALIENTES')] = 1
A[c('JALISCO','NAYARIT','DURANGO','COAHUILA DE ZARAGOZA','NUEVO LEON','TAMAULIPAS','SAN LUIS POTOSI','AGUASCALIENTES'),'ZACATECAS'] = 1
A['GUERRERO',c('MICHOACAN DE OCAMPO','PUEBLA','ESTADO DE MEXICO','MORELOS')] = 1
A[c('MICHOACAN DE OCAMPO','PUEBLA','ESTADO DE MEXICO','MORELOS'),'GUERRERO'] = 1
A['DURANGO',c('CHIHUAHUA','COAHUILA DE ZARAGOZA','ZACATECAS','NAYARIT','SINALOA')] = 1
A[c('CHIHUAHUA','COAHUILA DE ZARAGOZA','ZACATECAS','NAYARIT','SINALOA'),'DURANGO'] = 1
A['CHIHUAHUA',c('DURANGO','SINALOA','COAHUILA DE ZARAGOZA','SONORA')] = 1
A[c('DURANGO','SINALOA','COAHUILA DE ZARAGOZA','SONORA'),'CHIHUAHUA'] = 1
A['CHIAPAS',c('OAXACA','VERACRUZ LLAVE','TABASCO','CAMPECHE')] = 1
A[c('OAXACA','VERACRUZ LLAVE','TABASCO','CAMPECHE'),'CHIAPAS'] = 1
A['COLIMA',c('JALISCO','MICHOACAN DE OCAMPO')] = 1
A[c('JALISCO','MICHOACAN DE OCAMPO'),'COLIMA'] = 1　
A['COAHUILA DE ZARAGOZA',c('NUEVO LEON','ZACATECAS','DURANGO','CHIHUAHUA','SAN LUIS POTOSI')] = 1
A[c('NUEVO LEON','ZACATECAS','DURANGO','CHIHUAHUA','SAN LUIS POTOSI'),'COAHUILA DE ZARAGOZA'] = 1
A['SAN LUIS POTOSI',c('TABASCO','NUEVO LEON','COAHUILA DE ZARAGOZA','ZACATECAS','JALISCO','GUANAJUATO','QUERETARO DE ARTEAGA','HIDALGO','VERACRUZ LLAVE')] = 1
A[c('TABASCO','NUEVO LEON','COAHUILA DE ZARAGOZA','ZACATECAS','JALISCO','GUANAJUATO','QUERETARO DE ARTEAGA','HIDALGO','VERACRUZ LLAVE'),'SAN LUIS POTOSI'] = 1
A['CAMPECHE',c('TABASCO','QUINTANA ROO','YUCATAN')] = 1
A[c('TABASCO','QUINTANA ROO','YUCATAN'),'CAMPECHE'] = 1
A['BAJA CALIFORNIA SUR',c('BAJA CALIFORNIA','SONORA','SINALOA')] = 1
A[c('BAJA CALIFORNIA','SONORA','SINALOA'),'BAJA CALIFORNIA SUR'] = 1
A['BAJA CALIFORNIA',c('BAJA CALIFORNIA SUR','SONORA')] = 1
A[c('BAJA CALIFORNIA SUR','SONORA'),'BAJA CALIFORNIA'] = 1
A['AGUASCALIENTES',c('ZACATECAS','JALISCO')] = 1
A[c('ZACATECAS','JALISCO'),'AGUASCALIENTES'] = 1
A['YUCATAN',c('CAMPECHE','QUINTANA ROO')] = 1
A[c('CAMPECHE','QUINTANA ROO'),'YUCATAN'] = 1
A['TAMAULIPAS',c('NUEVO LEON','SAN LUIS POTOSI','VERACRUZ LLAVE')] = 1
A[c('NUEVO LEON','SAN LUIS POTOSI','VERACRUZ LLAVE'),'TAMAULIPAS'] = 1
A['TABASCO',c('CAMPECHE','CHIAPAS','VERACRUZ LLAVE')] = 1
A[c('CAMPECHE','CHIAPAS','VERACRUZ LLAVE'),'TABASCO'] = 1
A['SONORA',c('BAJA CALIFORNIA','BAJA CALIFORNIA SUR','SINALOA','CHIHUAHUA')] = 1
A[c('BAJA CALIFORNIA','BAJA CALIFORNIA SUR','SINALOA','CHIHUAHUA'),'SONORA'] = 1
A['SINALOA',c('SONORA','CHIHUAHUA','DURANGO','NAYARIT')] = 1
A[c('SONORA','CHIHUAHUA','DURANGO','NAYARIT'),'SINALOA'] = 1
A['QUINTANA ROO',c('YUCATAN','CAMPECHE')] = 1
A[c('YUCATAN','CAMPECHE'),'QUINTANA ROO'] =1
A['QUERETARO DE ARTEAGA',c('SAN LUIS POTOSI','HIDALGO','ESTADO DE MEXICO','GUANAJUATO','MICHOACAN DE OCAMPO')] = 1
A[c('SAN LUIS POTOSI','HIDALGO','ESTADO DE MEXICO','GUANAJUATO','MICHOACAN DE OCAMPO'),'QUERETARO DE ARTEAGA'] = 1
A['PUEBLA',c('ESTADO DE MEXICO','MORELOS','TLAXCALA','GUERRERO','OAXACA','VERACRUZ LLAVE','HIDALGO')] = 1
A[c('ESTADO DE MEXICO','MORELOS','TLAXCALA','GUERRERO','OAXACA','VERACRUZ LLAVE','HIDALGO'),'PUEBLA'] = 1
A['TLAXCALA',c('ESTADO DE MEXICO','HIDALGO','PUEBLA')] = 1
A[c('ESTADO DE MEXICO','HIDALGO','PUEBLA'),'TLAXCALA'] = 1
A['OAXACA',c('GUERRERO','PUEBLA','VERACRUZ LLAVE','CHIAPAS')] = 1
A[c('GUERRERO','PUEBLA','VERACRUZ LLAVE','CHIAPAS'),'OAXACA'] =1

```



```{r eval=FALSE, include=FALSE}

## Rescaling
inputs <- city_summary %>%
  mutate(
    market_state_city = (log(market_mean_city) - mean(log(market_mean_city))) /
      sd(log(market_mean_city)),
    
    income_state_city = (log(appraisal_mean_city) - mean(log(appraisal_mean_city))) /
      sd(log(appraisal_mean_city)),
    
    appraisal_state_city = (log(appraisal_mean_city) -
                            mean(log(appraisal_mean_city))) /
      sd(log(appraisal_mean_city)),
    
    mar_2_inc_city = (mar_2_inc_mean_city - mean(mar_2_inc_mean_city)) / sd(mar_2_inc_mean_city),
    
    app_2_inc_city = (app_2_inc_mean_city - mean(app_2_inc_mean_city)) / sd(app_2_inc_mean_city),
    
    mar_2_app_city = (mar_2_app_mean_city - mean(mar_2_app_mean_city)) / sd(mar_2_app_mean_city),
    
    age_city = (age_mean_city - mean(age_mean_city)) / sd(age_mean_city)) %>% 
  dplyr::select(
    market_state_city,
    income_state_city,
    appraisal_state_city,
    mar_2_inc_city,
    app_2_inc_city,
    mar_2_app_city,
    age_city,
    ID_state,
    n_city,
    sum_y,
    sum_y2
  )

```

```{r}
## Train / Test split
set.seed(1234)
model2 = stan_model('Model2.stan')

splited_inputs <- split(inputs, sample(rep(1:5, 176)))

for (i in 1:5){
  a <- c(1,2,3,4,5)[-i]
  inputs_test = splited_inputs[[i]]
  inputs_train = rbind(splited_inputs[[a[1]]] ,splited_inputs[[a[2]]],splited_inputs[[a[3]]],splited_inputs[[a[4]]]) 

  y_train = inputs_train$sum_y
  y_test = inputs_test$sum_y

## Inputs for STAN
  n_city_train = inputs_train$n_city
  n_city_test = inputs_test$n_city


  X_train = inputs_train %>% dplyr::select(-ID_state,-n_city,-sum_y, -sum_y2)
  X_test = inputs_test %>% dplyr::select(-ID_state,-n_city,-sum_y, -sum_y2)

  N_train = nrow(X_train)
  N_test = nrow(X_test)

  D = ncol(X_train)
  S = length(unique(data_sub$state))
  state_train = inputs_train$ID_state
  state_test = inputs_test$ID_state

  model2_data = list(S=S,N_train=N_train, N_test=N_test, D=D,
                   state_train = state_train,state_test = state_test,
                  X_train=X_train, X_test=X_test,W = A,W_n = sum(A) / 2,
                  n_city_train = n_city_train, 
                     n_city_test = n_city_test,
                     y = y_train)
  fit2 <- sampling(model2, data=model2_data,seed=1234)
  name = paste('model2_cv',as.character(i),'.rds',sep = "")
  saveRDS(fit2, file = name)
}

```




```{r eval=FALSE, include=FALSE}
MSE <- c()
for (i in 1:5){
  name = paste('model2_cv',as.character(i),'.rds',sep = "")
  fit1 <- readRDS(file = name)
  sims <- rstan::extract(fit1)
  y_hat <- apply(X = sims$y_rep_cv, MARGIN = 2, FUN = median)  
  
  test_df = data.frame(ID_state = inputs_test$ID_state,
                       y_test = y_test,
                       y_hat = y_hat)
  
  test_df <- test_df %>% 
    summarize(y_sum_test = sum(y_test),
              y_sum_hat = sum(y_hat)) %>% 
    arrange(desc(y_sum_test)) %>% 
    ungroup()
  
  mse_baseline = mean((test_df$y_sum_test) ** 2)
  mse = mean((test_df$y_sum_hat - test_df$y_sum_test) ** 2)
  MSE <- c(MSE,mse)
}
average_MSE <- mean(MSE)
SD_MSE <- sd(MSE)
```



```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
## Allocate train
y_train = inputs$sum_y

## Inputs for STAN
n_city_train = inputs$n_city
X_train = inputs %>% dplyr::select(-ID_state,-n_city,-sum_y, -sum_y2)

N_train = nrow(X_train)
D = ncol(X_train)
S = length(unique(data_sub$state))
state_train = inputs$ID_state
model2_data = list(S=S,N_train=N_train, D=D,
                   state_train = state_train,
                    X_train=X_train, W = A,W_n = sum(A) / 2,
                    n_city_train = n_city_train, 
                     y = y_train)
model2_full = stan_model('model2_full.stan')
fit2 <- sampling(model2_full, data=model2_data,seed=1234)
saveRDS(fit2, file = "model2_full.rds")
```


```{r}
## simulate the fake data
library(matlib)
set.seed(1234)
S = length(unique(data_sub$state))
n_city_train = inputs$n_city
X_train = inputs %>% dplyr::select(-ID_state,-n_city,-sum_y, -sum_y2)
N_train = nrow(X_train)
D = ncol(X_train)
state_train = inputs$ID_state
tau <- rgamma(n=1,shape = 2,rate = 2)
D <- diag(colSums(A))
W <- A
theta <- runif(n=1,min = 0,max = 1)
sigma <- tau * inv(D - W)
phi <- mvrnorm(n=1, mu = rep(0,32), Sigma = sigma )
beta <- rcauchy(n=7,location = 0,scale = 2.5)
alpha <- rcauchy(n=1,location = 0,scale = 10)
y <- c()
for (i in N_train){
  zero_d <- runif(n=1,min = 0,max = 1)
  
  if (zero_d < theta){
    new_y <- 0
  }else{
    prob <- invlogit(alpha + phi[state_train[i]] + X_train[i,] * beta)
    new_y <-rbinom(n=1,size = n_city,prob = prob)   
  }  
  y <- c(y,new_y)
}
fake_data = list(S=S,N_train=N_train, D=D,
                   state_train = state_train,
                    X_train=X_train, W = A,W_n = sum(A) / 2,
                    n_city_train = n_city_train, 
                     y = y)

## fit the model with the fake data
fake_fit <- sampling(model2_full, data=fake_data,seed=1234)

## test the recoverage
sim_fake <- rstan::extract(fake_fit)
true_phi <- phi
posterior_phi <- as.matrix(sim_fake, pars = c('phi'))
true_beta <- beta
posterior_beta <- as.matrix(sim_fake, pars = c('beta'))
true_alpha_tau_theta <- c(alpha,tau,theta)
posterior_alpha_tau_theta <- as.matrix(sim_fake, pars = c('alpha','tau','theta'))
```

```{r}
mcmc_recover_hist(posterior_phi, true = true_phi)
mcmc_recover_hist(posterior_beta, true = true_beta)
mcmc_recover_hist(posterior_alpha_tau_theta, true = true_alpha_tau_theta)
```





