---
title: "BDA_yi"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Imports, message=FALSE, warning=FALSE}
library(rstan)
library(tidyverse)
library(arm)
library(ggplot2)
library(gridExtra)
library(bayesplot)
library(shinystan)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

# Setting the preamble

# Data

```{r Load the data, message=FALSE, warning=FALSE}
file <-  '../data/core.txt'
data <- read_delim(file = file, delim = '|')

# Sample the data
pct = 1
# pct = 0.1
# pct = 0.01
set.seed(seed = 42)
sample_size = round(pct * nrow(data))
sample <- sample(x = nrow(data), size = sample_size, replace = F)
data = data[sample, ]

## Selecting the relevant columns for the analysis
data_sub <- data %>% dplyr::select(
  state,
  city,
  county,
  zip,
  asset_market_value,
  mar_2_app,
  appraisal_value,
  app_2_inc,
  client_income,
  mar_2_inc,
  age,
  sex_F,
  condition_U,
  y)

geo <- data_sub %>% group_by(city) %>% 
  summarize(market_mean = mean(asset_market_value),
            appraisal_mean = mean(appraisal_value),
            income_mean = mean(client_income),
            mar_2_inc_mean = mean(mar_2_inc),
            app_2_inc_mean = mean(app_2_inc),
            mar_2_app_mean = mean(mar_2_app),
            age_mean = mean(age),
            n_city = n(),
            y_sum = sum(y)) %>% ungroup()
state = c()
for (i in geo$city){state = c(state, unique(data$state[which(data$city==i)])[1])}
geo <- cbind(geo,state)
```

```{r Prep data for STAN model}
## Rescaling
inputs <- geo %>%
  mutate(
    income_st = (income_mean - mean(income_mean)) / sd(income_mean),
    appraisal_st = (appraisal_mean - mean(appraisal_mean)) / sd(appraisal_mean),
    market_st = (market_mean - mean(market_mean)) / sd(market_mean),
    mar_2_inc_st = (mar_2_inc_mean - mean(mar_2_inc_mean)) / sd(mar_2_inc_mean),
    app_2_inc_st = (app_2_inc_mean - mean(app_2_inc_mean)) / sd(app_2_inc_mean),
    mar_2_app_st = (mar_2_app_mean - mean(mar_2_app_mean)) / sd(mar_2_app_mean),
    age_st = (age_mean - mean(age_mean)) / sd(age_mean)
         ) %>% 
  dplyr::select(
    income_st,
    mar_2_inc_st,
    appraisal_st,
    app_2_inc_st,
    mar_2_app_st,
    market_st,
    age_st,
    state,
    number,
    y_sum
  )
```

```{r}
N = length(unique(inputs$state))
A <- matrix(0,ncol = N,nrow = N)
colnames(A) = unique(inputs$state)
rownames(A) = unique(inputs$state)
A['NUEVO LEON',c('ZACATECAS','SAN LUIS POTOSI','TAMAULIPAS','COAHUILA DE ZARAGOZA')] = 1
A[c('ZACATECAS','SAN LUIS POTOSI','TAMAULIPAS','COAHUILA DE ZARAGOZA'),'NUEVO LEON'] = 1
A['VERACRUZ LLAVE',c('TAMAULIPAS','SAN LUIS POTOSI','HIDALGO','PUEBLA','OAXACA','CHIAPAS','TABASCO')] = 1
A[c('TAMAULIPAS','SAN LUIS POTOSI','HIDALGO','PUEBLA','OAXACA','CHIAPAS','TABASCO'),'VERACRUZ LLAVE'] = 1
A['GUANAJUATO',c('MICHOACAN DE OCAMPO','JALISCO','SAN LUIS POTOSI','ZACATECAS','QUERETARO DE ARTEAGA')] = 1
A[c('MICHOACAN DE OCAMPO','JALISCO','SAN LUIS POTOSI','ZACATECAS','QUERETARO DE ARTEAGA'),'GUANAJUATO'] = 1
A['NAYARIT',c('SINALOA','DURANGO','ZACATECAS','JALISCO')] = 1
A[c('SINALOA','DURANGO','ZACATECAS','JALISCO'),'NAYARIT'] = 1
A['JALISCO',c('NAYARIT','ZACATECAS','COLIMA','MICHOACAN DE OCAMPO','GUANAJUATO','AGUASCALIENTES')] = 1
A[c('NAYARIT','ZACATECAS','COLIMA','MICHOACAN DE OCAMPO','GUANAJUATO','AGUASCALIENTES'),'JALISCO'] = 1
A['DISTRITO FEDERAL',c('ESTADO DE MEXICO','MORELOS')] = 1
A[c('ESTADO DE MEXICO','MORELOS'),'DISTRITO FEDERAL'] = 1
A['MORELOS',c('GUERRERO','PUEBLA','ESTADO DE MEXICO')] = 1
A[c('GUERRERO','PUEBLA','ESTADO DE MEXICO'),'MORELOS'] = 1
A['MICHOACAN DE OCAMPO',c('COAHUILA DE ZARAGOZA','ESTADO DE MEXICO','JALISCO','GUANAJUATO','GUERRERO','QUERETARO DE ARTEAGA')] = 1
A[c('COAHUILA DE ZARAGOZA','JALISCO','ESTADO DE MEXICO','GUANAJUATO','GUERRERO','QUERETARO DE ARTEAGA'),'MICHOACAN DE OCAMPO'] = 1
A['ESTADO DE MEXICO',c('HIDALGO','QUERETARO DE ARTEAGA','MICHOACAN DE OCAMPO','GUERRERO','PUEBLA','TLAXCALA','MORELOS')] = 1
A[c('HIDALGO','QUERETARO DE ARTEAGA','MICHOACAN DE OCAMPO','GUERRERO','PUEBLA','TLAXCALA','MORELOS'), 'ESTADO DE MEXICO'] = 1
A['HIDALGO',c('ESTADO DE MEXICO','QUERETARO DE ARTEAGA','SAN LUIS POTOSI','VERACRUZ LLAVE','PUEBLA','TLAXCALA')] = 1
A[c('ESTADO DE MEXICO','QUERETARO DE ARTEAGA','SAN LUIS POTOSI','VERACRUZ LLAVE','PUEBLA','TLAXCALA'),'HIDALGO'] = 1
A['ZACATECAS',c('JALISCO','NAYARIT','DURANGO','COAHUILA DE ZARAGOZA','NUEVO LEON','TAMAULIPAS','SAN LUIS POTOSI','AGUASCALIENTES')] = 1
A[c('JALISCO','NAYARIT','DURANGO','COAHUILA DE ZARAGOZA','NUEVO LEON','TAMAULIPAS','SAN LUIS POTOSI','AGUASCALIENTES'),'ZACATECAS'] = 1
A['GUERRERO',c('MICHOACAN DE OCAMPO','PUEBLA','ESTADO DE MEXICO','MORELOS')] = 1
A[c('MICHOACAN DE OCAMPO','PUEBLA','ESTADO DE MEXICO','MORELOS'),'GUERRERO'] = 1
A['DURANGO',c('CHIHUAHUA','COAHUILA DE ZARAGOZA','ZACATECAS','NAYARIT','SINALOA')] = 1
A[c('CHIHUAHUA','COAHUILA DE ZARAGOZA','ZACATECAS','NAYARIT','SINALOA'),'DURANGO'] = 1
A['CHIHUAHUA',c('DURANGO','SINALOA','COAHUILA DE ZARAGOZA','SONORA')] = 1
A[c('DURANGO','SINALOA','COAHUILA DE ZARAGOZA','SONORA'),'CHIHUAHUA'] = 1
A['CHIAPAS',c('OAXACA','VERACRUZ LLAVE','TABASCO','CAMPECHE')] = 1
A[c('OAXACA','VERACRUZ LLAVE','TABASCO','CAMPECHE'),'CHIAPAS'] = 1
A['COLIMA',c('JALISCO','MICHOACAN DE OCAMPO')] = 1
A[c('JALISCO','MICHOACAN DE OCAMPO'),'COLIMA'] = 1　
A['COAHUILA DE ZARAGOZA',c('NUEVO LEON','ZACATECAS','DURANGO','CHIHUAHUA','SAN LUIS POTOSI')] = 1
A[c('NUEVO LEON','ZACATECAS','DURANGO','CHIHUAHUA','SAN LUIS POTOSI'),'COAHUILA DE ZARAGOZA'] = 1
A['SAN LUIS POTOSI',c('TABASCO','NUEVO LEON','COAHUILA DE ZARAGOZA','ZACATECAS','JALISCO','GUANAJUATO','QUERETARO DE ARTEAGA','HIDALGO','VERACRUZ LLAVE')] = 1
A[c('TABASCO','NUEVO LEON','COAHUILA DE ZARAGOZA','ZACATECAS','JALISCO','GUANAJUATO','QUERETARO DE ARTEAGA','HIDALGO','VERACRUZ LLAVE'),'SAN LUIS POTOSI'] = 1
A['CAMPECHE',c('TABASCO','QUINTANA ROO','YUCATAN')] = 1
A[c('TABASCO','QUINTANA ROO','YUCATAN'),'CAMPECHE'] = 1
A['BAJA CALIFORNIA SUR',c('BAJA CALIFORNIA','SONORA','SINALOA')] = 1
A[c('BAJA CALIFORNIA','SONORA','SINALOA'),'BAJA CALIFORNIA SUR'] = 1
A['BAJA CALIFORNIA',c('BAJA CALIFORNIA SUR','SONORA')] = 1
A[c('BAJA CALIFORNIA SUR','SONORA'),'BAJA CALIFORNIA'] = 1
A['AGUASCALIENTES',c('ZACATECAS','JALISCO')] = 1
A[c('ZACATECAS','JALISCO'),'AGUASCALIENTES'] = 1
A['YUCATAN',c('CAMPECHE','QUINTANA ROO')] = 1
A[c('CAMPECHE','QUINTANA ROO'),'YUCATAN'] = 1
A['TAMAULIPAS',c('NUEVO LEON','SAN LUIS POTOSI','VERACRUZ LLAVE')] = 1
A[c('NUEVO LEON','SAN LUIS POTOSI','VERACRUZ LLAVE'),'TAMAULIPAS'] = 1
A['TABASCO',c('CAMPECHE','CHIAPAS','VERACRUZ LLAVE')] = 1
A[c('CAMPECHE','CHIAPAS','VERACRUZ LLAVE'),'TABASCO'] = 1
A['SONORA',c('BAJA CALIFORNIA','BAJA CALIFORNIA SUR','SINALOA','CHIHUAHUA')] = 1
A[c('BAJA CALIFORNIA','BAJA CALIFORNIA SUR','SINALOA','CHIHUAHUA'),'SONORA'] = 1
A['SINALOA',c('SONORA','CHIHUAHUA','DURANGO','NAYARIT')] = 1
A[c('SONORA','CHIHUAHUA','DURANGO','NAYARIT'),'SINALOA'] = 1
A['QUINTANA ROO',c('YUCATAN','CAMPECHE')] = 1
A[c('YUCATAN','CAMPECHE'),'QUINTANA ROO'] =1
A['QUERETARO DE ARTEAGA',c('SAN LUIS POTOSI','HIDALGO','ESTADO DE MEXICO','GUANAJUATO','MICHOACAN DE OCAMPO')] = 1
A[c('SAN LUIS POTOSI','HIDALGO','ESTADO DE MEXICO','GUANAJUATO','MICHOACAN DE OCAMPO'),'QUERETARO DE ARTEAGA'] = 1
A['PUEBLA',c('ESTADO DE MEXICO','MORELOS','TLAXCALA','GUERRERO','OAXACA','VERACRUZ LLAVE','HIDALGO')] = 1
A[c('ESTADO DE MEXICO','MORELOS','TLAXCALA','GUERRERO','OAXACA','VERACRUZ LLAVE','HIDALGO'),'PUEBLA'] = 1
A['TLAXCALA',c('ESTADO DE MEXICO','HIDALGO','PUEBLA')] = 1
A[c('ESTADO DE MEXICO','HIDALGO','PUEBLA'),'TLAXCALA'] = 1
A['OAXACA',c('GUERRERO','PUEBLA','VERACRUZ LLAVE','CHIAPAS')] = 1
A[c('GUERRERO','PUEBLA','VERACRUZ LLAVE','CHIAPAS'),'OAXACA'] =1
```





```{r}
## Inputs for STAN
y = inputs$y_sum
number = inputs$number
X = inputs %>% dplyr::select(-y_sum, -state,-number)
X = cbind(1,X)
colnames(X) <- c('intercept','income_st','mar_2_inc_st','appraisal_st','app_2_inc_st','mar_2_app_st','market_st','age_st')
n = nrow(X)
p = ncol(X)
state = as.vector(inputs$state)
unique_s = unique(state)
for (i in 1:32){
  s = as.character(unique_s[i])
  state[state==s] = as.numeric(i)
}
state = as.numeric(state)
```

### Baseline model

1. Modeling
```{r warning=FALSE}
## baseline model
baseline_data = list(N=n,p=p, number = inputs$number,X=X,y=y)
baseline_model =  stan_model('./yi_stan/yi_baseline_city.stan')
fit <- sampling(baseline_model, data = baseline_data,show_messages = FALSE,seed=1234)
print(fit,pars=c('beta','lp__'))
ss_mod = as.shinystan(fit)
launch_shinystan(ss_mod)

```


2. PPCs


### CAR model
1. Modeling

```{r}
## CAR model
CAR_data <- list(k = nrow(X),         # number of observations
               n = nrow(A),
               p = ncol(X),         # number of coefficients
               X = X,               # design matrix
               y = y,               # observed number of cases
               W = A,
               number = inputs$number,
               W_n = sum(A) / 2,    # number of neighbor pairs
               state = state)               # adjacency matrix
CAR_model = model <- stan_model('./yi_stan/yi_CAR_city.stan')
fit_car <- sampling(CAR_model, data = CAR_data,show_messages = FALSE,seed=1234)
print(fit_car,pars=c('phi','lp__'))
ss_mod_car = as.shinystan(fit_car)
launch_shinystan(ss_mod_car)
```
2. PPCs



### Hier model
1. Modeling

```{r}
## Hier model
Hier_data <- list(n = n,         # number of observations
               s = 32,
               p = p,         # number of coefficients
               X = X,               # design matrix
               y = y,               # observed number of cases
               number = inputs$number,
               state = state)               # adjacency matrix
Hier_model = model <- stan_model('./yi_stan/yi_hier_city.stan')
Hier_fit <- sampling(Hier_model, data = Hier_data,show_messages = FALSE,seed=1234)
print(Hier_fit,pars=c('alpha','alpha_s','lp__'))
Hier_ss_mod = as.shinystan(Hier_fit)
launch_shinystan(Hier_ss_mod)
```
2. PPCs

### extension: zero inflated possion binomial-mixture model

```{r}
## zero inflated model
zero_data <- list(k = nrow(X),         # number of observations
               n = nrow(A),
               p = ncol(X),         # number of coefficients
               X = X,               # design matrix
               y = y,               # observed number of cases
               W = A,
               number = inputs$number,
               W_n = sum(A) / 2,    # number of neighbor pairs
               state = state)               # adjacency matrix
zero_model <- stan_model('./yi_stan/zero-inflated-possion-binormal-mix.stan')
fit_zero <- sampling(zero_model, data = zero_data,show_messages = FALSE,seed=1234)
print(fit_zero,pars=c('phi','lp__'))
ss_mod_zero = as.shinystan(fit_zero)
launch_shinystan(ss_mod_zero)
```



## Covariates

[...] Put the list of covariates in the 

## Exploratory Data Analysis

[...] Show general relationships in the data

## Data Preprocessing

[...] Add standarization procedures

# Baseline Model

## Specification

## Results

## Evaluation

# Model Extension 1

## Specification

## Results

## Evaluation

# Model Extension 2

## Specification

## Results

## Evaluation

# Conclusion

# Stan Code







#### 12/1

```{r}
library(rstan)
library(tidyverse)
library(arm)
library(ggplot2)
library(gridExtra)
library(bayesplot)
library(shinystan)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
file <-  '../data/core.txt'
data <- read_delim(file = file, delim = '|')
# Sample the data
pct = 1
# pct = 0.1
# pct = 0.01
set.seed(seed = 42)
sample_size = round(pct * nrow(data))
sample <- sample(x = nrow(data), size = sample_size, replace = F)
data = data[sample, ]
## Selecting the relevant columns for the analysis
data_sub <- data %>% dplyr::select(
  state,
  city,
  county,
  zip,
  asset_market_value,
  mar_2_app,
  appraisal_value,
  app_2_inc,
  client_income,
  mar_2_inc,
  age,
  sex_F,
  condition_U,
  y,
  y2)
summary(data_sub)


## Group data by state and define the IDs
state_summary <- data_sub %>% 
  dplyr::select(state, 
                client_income, 
                appraisal_value,
                asset_market_value) %>% 
  group_by(state) %>% 
  summarize(n_state = n(),
            income_mean_state = mean(client_income),
            appraisal_mean_state = mean(appraisal_value),
            market_mean_state = mean(asset_market_value)) %>% 
  arrange(desc(n_state)) %>% 
  ungroup()
state_summary$ID_state = seq.int(nrow(state_summary))


## Group data by city and define the IDs
city_summary <- data_sub %>% 
  dplyr::select(city,
                state,
                client_income,
                appraisal_value,
                asset_market_value,
                mar_2_inc,
                mar_2_app,
                app_2_inc,
                age,
                y,
                y2) %>% 
  group_by(city, state) %>% 
  summarize(n_city = n(),
            income_mean_city = mean(client_income),
            appraisal_mean_city = mean(appraisal_value),
            market_mean_city = mean(asset_market_value),
            
            mar_2_inc_mean_city = mean(mar_2_inc),
            mar_2_app_mean_city = mean(mar_2_app),
            app_2_inc_mean_city = mean(app_2_inc),
            
            age_mean_city = mean(age),
            sum_y = sum(y),
            sum_y2 = sum(y2)) %>% 
  arrange(desc(n_city)) %>% 
  ungroup()


## Merge back into data
city_summary <- city_summary %>% 
  inner_join(y = state_summary[c('ID_state', 'state')], by = 'state')


## Rescaling
inputs <- city_summary %>%
  mutate(
    market_state_city = (log(market_mean_city) - mean(log(market_mean_city))) /
      sd(log(market_mean_city)),
    
    income_state_city = (log(appraisal_mean_city) - mean(log(appraisal_mean_city))) /
      sd(log(appraisal_mean_city)),
    
    appraisal_state_city = (log(appraisal_mean_city) -
                            mean(log(appraisal_mean_city))) /
      sd(log(appraisal_mean_city)),
    
    mar_2_inc_city = (mar_2_inc_mean_city - mean(mar_2_inc_mean_city)) / sd(mar_2_inc_mean_city),
    
    app_2_inc_city = (app_2_inc_mean_city - mean(app_2_inc_mean_city)) / sd(app_2_inc_mean_city),
    
    mar_2_app_city = (mar_2_app_mean_city - mean(mar_2_app_mean_city)) / sd(mar_2_app_mean_city),
    
    age_city = (age_mean_city - mean(age_mean_city)) / sd(age_mean_city)) %>% 
  dplyr::select(
    market_state_city,
    income_state_city,
    appraisal_state_city,
    mar_2_inc_city,
    app_2_inc_city,
    mar_2_app_city,
    age_city,
    ID_state,
    n_city,
    sum_y,
    sum_y2
  )


## Train / Test split
set.seed(seed = 81989843)
pct_train = 0.8
sample_size = round(pct_train * nrow(inputs))
sample <- sample(x = nrow(inputs), size = sample_size, replace = F)

## Allocate train
y = inputs$sum_y
y2 = inputs$sum_y2

inputs_train = inputs[sample, ]
y_train = y[sample]

## Allocate test
inputs_test = inputs[-sample, ]
y_test = y[-sample]

## Inputs for STAN
X_train = inputs_train %>% dplyr::select(-sum_y, -sum_y2)
X_test = inputs_test %>% dplyr::select(-sum_y, -sum_y2)

N_train = nrow(X_train)
N_test = nrow(X_test)

n_city_train = X_train$n_city
n_city_test = X_test$n_city

D = ncol(X_train)
S = length(unique(data_sub$state))
state = data_sub$ID_state[sample]

baseline_data = list(N_train=N_train, N_test=N_test, D=D, 
                     X_train=X_train, X_test=X_test,
                     n_city_train = n_city_train, 
                     n_city_test = n_city_test,
                     y_train = y_train)
```

```{r}
baseline_fit <- stan_model('./yi_stan/yi_baseline_city.stan')
fit <- sampling(baseline_fit, data = baseline_data,show_messages = FALSE,seed=1234)
```

```{r}
sims <- rstan::extract(fit)
# a_median <- median(sims$a)
# beta_median <- apply(X = sims$beta, MARGIN = 2, FUN = median)
# 
# y_hat <- N_test * invlogit(a_median + as.matrix(X_test)%*%beta_median)
# mode <- function(x) {
#   ux <- unique(x)
#   ux[which.max(tabulate(match(x, ux)))]
# }
y_hat <- apply(X = sims$y_rep_cv, MARGIN = 2, FUN = median)  
test_df = data.frame(ID_state = inputs_test$ID_state,
                     y_test = y_test,
                     y_hat = y_hat)
test_df <- test_df %>% 
  summarize(y_sum_test = sum(y_test),
            y_sum_hat = sum(y_hat)) %>% 
  arrange(desc(y_sum_test)) %>% 
  ungroup()
accuracy_baseline = mean(abs(test_df$y_sum_test) ** 2)
accuracy = mean(abs(test_df$y_sum_hat - test_df$y_sum_test) ** 2)
cat('\nBaseline MSE: ', accuracy_baseline * 100)
cat('\nLogistic MSE: ', accuracy * 100)
```


